
***Note*:** *This problems make uses of the element-to-element property of operations on numeric vectors to calculate pattern summation.* 

***How to do it?***: 


- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Sample Codes***: You can find similar codes for this assignment here [link](fa2020_assignment2_sample_code.html)


***Submissions***: Submit the link on Github of the assignment to Blackboard under Assignment 2. 


Problems
-------
-------


1. Calculate: 


$$
S = 1 + 4 + 7 + 11 +... + 100
$$
```{r}
x <- seq(1,101,3)
sum(x)
```

2. Calculate: 


$$
S = 1^{2020} + 2^{2020} + 3^{2020} +...+2020^{2020}
$$
```{r}
s <- c(1:2020)
sum(s^2020)
```

3. Calculate: 

$$
S = \frac{1}{1\cdot 2}+\frac{1}{2 \cdot 3}+...+\frac{1}{2020\cdot 2021}
$$
```{r}
a <- c(1:2020)
b <- c(2:2021)
sum(1/(a*b))
```

4. Calculate: 

$$
S = \frac{1}{1\cdot 2 \cdot 3}+\frac{1}{2 \cdot 3 \cdot 4}+...+\frac{1}{2020\cdot 2021 \cdot 2022}
$$
```{r}
l <- c(1:2020)
m <- c(2:2021)
n <- c(3:2022)
sum(1/(l*m*n))
```

5. Calculate: 

$$
S = 1^1 + 3^3 + 5^5 + 7^7 +....+2021^{2021}
$$
```{r}
e <- seq(1,2021,2)
sum(e^e)
```
