
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 6: Visualization - Aesthetic Mapping"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment6.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. 

```{r}
library(tidyverse)
spotify=read_csv('data.csv')
str(spotify)
```
```{r}
library(lubridate)
spotify$release_date <- ymd(spotify$release_date)
class(spotify$release_date)
```
```{r}
#make ms into s
spotify %>% mutate(duration_s= duration_ms/1000)
```

-------

1. For one continuous variable:

- Density

- Histogram 

- Boxplot

```{r}
library(ggplot2)
spotify %>% mutate(duration_s= duration_ms/1000) %>% 
ggplot()+ aes(x=duration_s) + geom_density()

#seeing where the density is the highest, most of the songs are in the same duration
```
 

```{r}
library(ggplot2)
spotify %>% mutate(duration_s= duration_ms/1000) %>% 
ggplot()+ aes(x=duration_s) + geom_histogram()

#very similar to density plot
```
```{r}
spotify %>% mutate(duration_s= duration_ms/1000) %>% filter(year>2019) %>% ggplot()+
  geom_boxplot(mapping = aes(x = duration_s))
#this is clear that many of the songs are in the 180-250s range for length with several outliars. 
```

-------

2. For one categorical variable


- Barplot
```{r}
ggplot(spotify, aes(x=explicit)) + geom_bar()
#majority of the songs are not explicit
```


-------

3. For one two continuous variables

- Scatter Plot
```{r}
spotify %>% filter(year==(2019)) %>% ggplot()+ aes(x=loudness,y=danceability) + geom_point()
#the cluster of loud music also ranks high in dancablity
```


- Line Plot
```{r}
spotify %>% filter(year>2019) %>% ggplot()+
  geom_line(mapping = aes(x=tempo,y=danceability))

#although scattered, the general trend is higher the tempo, higher the danceability

```


- Smooth Plot
```{r}
spotify %>% filter(year>2019) %>% ggplot()+
  geom_smooth(mapping = aes(x=danceability,y=tempo))
#same conclusion as above, just easier to determine
```

-------

4. For one continuous + one categorical variables

- Density
```{r}
spotify %>% filter(year>2015) %>% ggplot()+
  geom_density(mapping = aes(color=factor(year),x=duration_ms))

#as the years have gone on seems that the songs are getting longer
```
```{r}
spotify %>% filter(year>2000)
```


- BoxPlot

-------

5. For two categorical variables: barplot
```{r}
spotify %>% filter(year>2015) %>% ggplot()+
  geom_bar(mapping = aes(color=factor(explicit),x=danceability))

#the explicit songs are skewed towards higher danceability
```



