
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 10: Predictive Modeling - Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```








-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes
str(df)
```
```{r}
sapply(df, function(x) sum(is.na(x)))
#no missing values
```

- Set seed to be 2020. 
```{r}
library(caret)
names(df)[9]<-'target'

set.seed(2020)
```

- Partition the data into 80% training and 20% testing.  
```{r}
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

-------

2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
```{r}
library(rpart) #load the rpart package
# Create a tree
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
```
  
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
library(e1071)
pred <- predict(tree_model, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```
  
  
  - Plot the tree
```{r}
library(rattle)
fancyRpartPlot(tree_model)
```
  
  
  - Plot the variable importance by the tree
```{r}
barplot(tree_model$variable.importance)
```

-------

3. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
```{r}
library(randomForest)
forest_model = randomForest(target ~ ., data=df_train, ntree = 1000)

```

  
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
pred <- predict(forest_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```
  
  - Plot the variable importance by the forest
```{r}
importance(forest_model)
```
  

-------

4. Compare the accuracy of a forest of 1000 trees and a forest of 2000 trees. 
```{r}
forest_model2 <- randomForest(target ~.,data=df_train,ntree=2000)
pred <- predict(forest_model2, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```

-------

5. Using Caret, create a tree with maximum depth of 3 and forest of 1000 trees. Compare the accuracy of these two models.

```{r}
model1 <- train(target~., data=df_train, 
                method = "rpart2",
                maxdepth="3") 
pred <- predict(model1, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```
```{r}
model2 <- train(target~., data=df_train, 
                method = "rf",
                ntree = 1000) 
pred <- predict(model2, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```


-------

6. Plot variable importance by the two models in 5. 
```{r}
varImp(model1)
```
```{r}
varImp(model2)
```

