
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(tidyverse)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df2 <- df %>% filter(Date_reported==Sys.Date()) %>% arrange(-Cumulative_deaths)
```



2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths

```{r}
library(gganimate)
d1 <- df %>% group_by(Date_reported, Country) %>% summarise(Cumulative_deaths)
d2 <- d1 %>% mutate(rank=rank(-Cumulative_deaths)) %>% ungroup()
a1 <- d2 %>% filter(rank <= 3) %>% 
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
animate(a1,nframes = 150)
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 

