
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment12.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes
```

- Set seed to be 2020. 
- The target variable is `diabetes`
- Partition the data into 80% training and 20% testing. 
```{r}
library(caret)
names(df)[9]<-'target'

set.seed(2020)

splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

-------

2. Use cross-validation of 30 folds to tune random forest (method='rf').  What is the `mtry` value that produces the greatest accuracy?


```{r}

tuneGrid = expand.grid(mtry = 2:4)


trControl = trainControl(method = "cv",
                         number = 30)

forest_cv <- train(target~., data=df_train, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
```

```{r}
plot(forest_cv)
```

 
-------

```{r}
install.packages("ranger")
```

3. Use cross-validation with of 30 folds to tune random forest (method='ranger').  What are the parameters that produce the greatest accuracy?

```{r}
tuneGrid = expand.grid(mtry = 2:4)


trControl = trainControl(method = "cv",
                         number = 30)

forest_cv <- train(target~., data=df_train, 
                                method = "ranger", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
```

-------

4. Go to https://topepo.github.io/caret/available-models.html and pick a classification model.  Tune the classification model using cross-validation of 30 folds. 

5. Compare the three models in question 2, 3, and 4 to select the final model.  Evaluate the accuracy of the final model on the test data. 
