---
title: "Assignment 3"
author: "Kathryn Doorley"
date: "September 2, 2020"
---
***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard under Assignment 3. 

-------

Problems
-------
 \

1. Use `read.csv` to import the Covid19 data from WHO: https://covid19.who.int/WHO-COVID-19-global-data.csv
```{r}
df <- read.csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
```


2. How many columns and rows the data have? How many categorical variables the data have? 
```{r}
#39001 rows and 8 cols
str(df)
```


3. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
colSums(is.na(df))
```
```{r}
library(lubridate)
```


4. Correct the types of the variables that are mismatched type.
```{r}
str(df)


```


5. Find the daily average number of cases globally.
```{r}
mean(df$New_cases)
```


6. Find the maximum cases a day.  
```{r}
max(df$New_cases)
```

7. Which country at which date has the maximum number of cases a day? *Hint:  use the `which.max` function to find the index of the row that contains the maximum number of cases. Then from the row index, find that country and the date. *
```{r}
which.max(df$New_cases)
df[16888,]
```

8.  How many WHO's region there are in the data?
```{r}
table(df$WHO_region)
```


9.  Find the daily average cases in EUROPE. Find the daily average cases by regions.
```{r}
d1=df[(df$WHO_region=="EURO"),]
mean(d1$New_cases)

```


10. Find the daily average number of cases in the US in the last 5 days. 

```{r}
d1 = df[(Sys.Date()- df$date_reported<7),]
mean(d1)
```

11. Which day of the week has the most number of cases in South East Asia (SEARO region)?
```{r}
d1=df[(df$WHO_region=="SEARO"),]
max(d1$New_cases)
df[78761,]
```

