
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Fall 2020 - Math 421 - Midterm"
---

-------

# Instruction

The midterm has two components: the Rmarkdown notebook (html) and the presentation.  The notebook and the presentation file are due Nov 26. We will do the presentation in class 

**The notebook:** The notebook should be created using `rmarkdown` (like other assignments). The notebook should have a title. 


**The Presentation:** Present your results in 5-10 minutes. To make the presentation using Rmarkdown, do the follows: 

    - In Rstudio -> File -> New File -> R markdown
    
    - In the left panel, click to Presentation -> Click OK
    
    - Now you have an Rmarkdown that can be knitted to be a html presentation 
    

**Data:**  The data for the mid-term project is the Rhode Island Department of Health Hospital Discharge Data.  Each row of the data presents a patient. 

Link: https://drive.google.com/open?id=15QNBf6YYKocK2nNIfpKDer58kQnCPNZJ 

-------

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data. 

```{r}
library(haven)
mid_df <- read_sas('hdd0318cy.sas7bdat')
```

    
3. Filter the data to have only patients of the year 2018 (`yod==2018`)

```{r}
midterm_2018 <- mid_df %>% filter(yod==2018)
```

    
4. Select to work with only following variables: 

```{r, eval=FALSE}
                      "yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day"
```
 

*Notice*:  You may want to save the current data on your laptop if you want to read it again.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

5. What are variables that have missing values?
 
6. Remove all variables with missing values
 
7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?

8. Which month admitted the most number of patients? Which month admitted the most number of male patients?

9. Which month has the most number of teenage female patients?

10. Which provider has the most number of female patients in October? 

11. Is female patients older than male patients, on average? 

12. Calculate the average age of patients by months. Which month has the oldest patients on average age?

13. What is the name of the provider that has the highest total charge?

14. What is the name of the provider that has the least total charge for teenage male?

15. Calculate the length of stays by races.  Which race has the longest length of stays on average?

16. On average, how much a 20 year-old male white get charged for staying 1 day?

-------

## II. Data Visualization

Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis

2. Make an animation

-------

## III. Predictive Models

Continue with the data from part I. Use the follows as the target and input variables: 

*Target Variable*: Create the target variable taking value of 

  - `low cost` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high cost` otherwise. 

*Input Variables*:

  - "age","sex","raceethn","provider","moa","mod","admtype","campus", "er_mode", 'los'
  
-------

1. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 

3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 3 different models (i.e. three different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model. 

4. Plot the comparison of the models in 2. 

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 

6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 

-------